brados: declarative, programmable object storage

### 2016-05-25 paper.tex
addresing some of peters feedback
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

% why is it interesting
The standardization of the POSIX file I/O interface has been a major success,
allowing application developers to avoid vendor lock-in. However, large-scale
storage systems have been dominated by proprietary products, preventing
exploration of alternative interfaces and complicating future migration paths,
eliminating the benefits of commodity systems. But the recent availability of
high-performance open-source storage systems is changing this because these
systems are modifiable, enabling interface change, and reducing the risks of
lock-in. The widely deployed Ceph distributed storage system is an example of
a storage system that supports application-specific extensions in the form of
custom I/O interfaces to objects managed by the underlying RADOS object
storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
reliant upon these extensions as is shown in Figure~\ref{fig:objclass-dev} by
a marked increase in the number of object operations that are packaged as part
of the Ceph distribution and widely used by internal Ceph subsystems and by
applications such as OpenStack Swift and Cinder~\cite{openstack}.

@@

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerns that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

@@

Figure~\ref{fig:ceph} illustrates the collection of components commonly
referred to as Ceph. At the bottom, a cluster of 10s--10,000s \emph{object
storage devices} compose the distributed object storage system called RADOS.
Widely deployed applications such as the S3/Swift--compliant RADOS Gateway
(RGW), RADOS Block Device (RBD), and the POSIX Ceph File System are built upon
the \emph{librados} client layer that presents a fault-tolerant always-on view
of the RADOS cluster.

@@

\section{Physical Design Space}

@@

As Ceph continues to develop into a target for building distributed services
and application, it is important to not restrict use to developers seasonsed
in distributed programming and those with knowledge of intracacies of the Ceph
performance model. To this end we present Brados, a declarative language with
relational semantics for defining the behavior of RADOS object interfaces.
Brados uses a simple declarative data model and can reduce the learning curve
for new users, and allows existing developers to increase productivity by
writing less code that is more portable.

The Brados language corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence’’ offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage decide
classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.

@@

  scratch :write_op, [:type, :pos, :epoch] => [:data]
  scratch :read_op,  [:type, :pos, :epoch] => [:data]
  scratch :trim_op,  [:type, :pos, :epoch] => [:data]
  scratch :fill_op,  [:type, :pos, :epoch] => [:data]
  scratch :seal_op,  [:type, :pos, :epoch] => [:data]

  # op did or did not pass the epoch guard
  scratch :valid_op,   [:type, :pos, :epoch] => [:data]
  scratch :invalid_op, [:type, :pos, :epoch] => [:data]

  # op's position was or was not found in the log
  scratch :found_op,    [:type, :pos, :epoch] => [:data]
  scratch :notfound_op, [:type, :pos, :epoch] => [:data]
end

@@

The write and fill interfaces are implemented similarily, and are show in
Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
write collection is created if the operation position is not found in the log.
A valid write operation is then merged into log, otherwise a read only error is
returend indicating that hte log position was already written to.
The fill operation is identical except a the fill state is set on the log entry.

@@

>>>>

% why is it interesting
The standardization of the POSIX file I/O interface has been a major success,
allowing application developers to avoid vendor lock-in. However, large-scale
storage systems have been dominated by proprietary products, preventing
exploration of alternative interfaces and complicating future migration paths,
eliminating the benefits of commodity systems. But the recent availability of
high-performance open-source storage systems is changing the way people
approach storage because these
systems are modifiable, enabling interface change and reducing the risks of
lock-in. The widely deployed Ceph distributed storage system is an example of
a storage system that supports application-specific extensions in the form of
custom I/O interfaces to objects managed by the underlying RADOS object
storage system~\cite{weil:osdi06,weil:pdsw07}. Organizations are increasingly
reliant upon these extensions. Figure~\ref{fig:objclass-dev} shows
a marked increase in the number of object operations that are packaged as part
of the Ceph distribution and widely used by internal Ceph subsystems and by
applications such as OpenStack Swift and Cinder~\cite{openstack}.

@@

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that
``future-proofs'' an
implementation against hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerns that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

@@

Figure~\ref{fig:ceph} illustrates the collection of components commonly
referred to as Ceph. At the bottom, a cluster of 10s--10,000s of \emph{object
storage devices} compose the distributed object storage system called RADOS.
Widely deployed applications such as the S3/Swift--compliant RADOS Gateway
(RGW), RADOS Block Device (RBD), and the POSIX Ceph File System are built upon
the \emph{librados} client layer that presents a fault-tolerant always-on view
of the RADOS cluster.

@@

\section{Physical Design Space}
\label{sec:pd}

@@

As Ceph continues to develop into a target for building distributed services
and application, it is important to not restrict use to developers seasoned
in distributed programming and those with knowledge of intricacies of the Ceph
performance model. To this end we present Brados, a declarative language with
relational semantics for defining the behavior of RADOS object interfaces.
Brados uses a simple declarative data model and can reduce the learning curve
for new users, and allows existing developers to increase productivity by
writing less code that is more portable.

The Brados language corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence'' offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage decide
classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.

@@

  scratch :write_op, op.schema
  scratch :read_op,  op.schema
  scratch :trim_op,  op.schema
  scratch :fill_op,  op.schema
  scratch :seal_op,  op.schema

  # op did or did not pass the epoch guard
  scratch :valid_op,   op.schema
  scratch :invalid_op, op.schema

  # op's position was or was not found in the log
  scratch :found_op,    op.schema
  scratch :notfound_op, op.schema
end

@@

The write and fill interfaces are implemented similarily, and are show in
Listing~\ref{lst:write} and Listing~\ref{lst:fill}, respectively.  A valid
write collection is created if the operation position is not found in the log.
A valid write operation is then merged into log, otherwise a read only error is
returend indicating that the log position was already written to.
The fill operation is identical except a the fill state is set on the log entry.

@@

Amazingly these few code snippets express the semantics of the entire storage
device interface requirements in CORFU. For reference the ZLog implementation
of the CORFU storage interface in C++ comprises nearly 700 lines of code and
was designed to store log entries and metadata in the key-value interface,
thus requiring a lengthy rewrite to realize the performance advantage
available by using the bytestream interface, as was highlighted in
Section~\ref{sec:pd}. Furthermore, the complexity introduced by using the
bytestream interface would grow the amount of code written in the optimized
version.

But beyond the convience of writing less code, it is far easier for the
programmer writing an interface such as CORFU to convinence herself of the
correctness of the high-level details of the implementation without being
distracted by issues related to physical design or the many other gotchas that
one must deal with when writing low-level systems software.
